---
title: "Challenge 7: Incorporating Multiple Inputs"
subtitle: "Functions + Fish"
author: "Zack Kramer"
format: html
editor: source
embed-resources: true
code-tools: true
toc: true
code-fold: true
execute: 
  error: true
  echo: true
  message: false
  warning: false
---

```{r}
#| label: setup

library(tidyverse)

```

```{r}
#| label: import-data

fish <- read_csv(here::here('data','BlackfootFish.csv'))

```


## Condition Index

A frequently used measurement for fish health is a condition index [(Wikipedia article)](https://en.wikipedia.org/wiki/Condition_index_in_fish). The following
simple equation can be used to calculate the approximate condition index of a
fish:

$$\text{condition index} = \frac{weight}{length^3} \times 100$$

**1. There are specific units required for the calculation of a condition index -- length must be in centimeters and weight must be in grams. The weight data for the Blackfoot River fish were collected in grams; the length data were collected in millimeters. Transform the length data to the correct units.**

```{r}
#| label: transform-data-to-have-correct-units

fish <- fish |>
  mutate(length = length/10)


```


**2. Collecting data of this sort can be very messy! Write a function that will replace unlikely `length` and `weight` measurements with `NA`. Your function should accept at least three inputs:**

+ **a vector of measurements,**
+ **the minimum reasonable value,**
+ **the maximum reasonable value.**

**If a value falls outside these bounds, you should replace it with an `NA`.**

::: callout-tip
If you are struggling with the structure of your function, I would suggest reading the [Mutating Function](https://r4ds.hadley.nz/functions.html#mutate-functions) from R4DS.
:::

```{r}
#| label: function-to-change-impossible-measurements-to-NA

replace_unlikely_vals <- function(x, x_min, x_max){
  
  if (!is.numeric(x) | length(x) <= 1) {
    stop("Error: Input vector must be numeric with a length not greater than 1.")
  }
  
  case_when(
    x < x_min ~ NA,
    x > x_max ~ NA,
    .default = x
  )
  
}


```


**3. After consulting the [Montana Record Table](https://fwp.mt.gov/fish/anglingData/records) for the four species of trout included in these data, I have conjectured that it is unlikely to have measurements for fish below 5 cm and above 80 cm in length or below 10 g and above 4,000 g in weight. Use your function to modify the `length` and `weight` columns of the `fish` dataset based on my cutoffs.**

```{r}
#| label: modify-impossible-values-of-length-weight

fish <- fish |>
  mutate(length = replace_unlikely_vals(length, 5, 80),
         weight = replace_unlikely_vals(weight, 10, 4000))

```


**3. Write a function to calculate the condition index of a fish, given inputs of weight and length.**

::: callout-warning
Consider whether you will write a function that works with vectors (accepting
vectors as inputs) or a function that works with data frames (accepting a 
dataset and variable names as inputs)!
:::

```{r}
#| label: condition-function

calculate_condition_index <- function(df, len, wt){
  
  df <- df |>
    mutate(condition_index = ( {{wt}} / ( {{len}} * {{len}} * {{len}} ) * 100))
  
  return(df)
    
}

```

**4. Make a thoughtful visualization of how fish conditions have varied over the duration of this study.**

::: callout-warning
### To be thoughtful...

... you need to use at least a third variable!
:::

```{r}
#| label: condition-indices-over-time

calculate_condition_index(fish, length, weight) |>
  group_by(year, species) |>
  summarize(median_condition = median(condition_index, na.rm = TRUE)) |>
  ggplot(mapping = aes(x = year, 
                       y = median_condition)) + 
  geom_col() + 
  labs(x = "Year",
       y = "",
       title = "Median Fish Condition Index from 1989 to 2006") + 
  facet_wrap(~ species) + 
  theme_light() + 
  scale_x_continuous(breaks = seq(1990, 2006, by = 2))
  
```
